\section*{Acknowledgments}

%% TODO: Michael needs to check what parts of these disclaimers will actually be needed and what we can remove
% * Other names and brands are the property of their respective owners.

Software and workloads used in performance tests may have been
optimized for performance only on Intel microprocessors.  Performance
tests, such as SYSmark and MobileMark, are measured using specific
computer systems, components, software, operations and functions.  Any
change to any of those factors may cause the results to vary.  You
should consult other information and performance tests to assist you
in fully evaluating your contemplated purchases, including the
performance of that product when combined with other products.  For
more information go to \url{http://www.intel.com/performance}.

Intel's compilers may or may not optimize to the same degree for
non-Intel microprocessors for optimizations that are not unique to
Intel microprocessors. These optimizations include SSE2, SSE3, and
SSSE3 instruction sets and other optimizations. Intel does not
guarantee the availability, functionality, or effectiveness of any
optimization on microprocessors not manufactured by
Intel. Microprocessor-dependent optimizations in this product are
intended for use with Intel microprocessors. Certain optimizations not
specific to Intel microarchitecture are reserved for Intel
microprocessors. Please refer to the applicable product User and
Reference Guides for more information regarding the specific
instruction sets covered by this notice.

% usual LLNL boilerplate
This work was performed under the auspices of the U.S. Department of Energy by
Lawrence Livermore National Laboratory under Contract DE-AC52-07NA27344.

% Stephen: Must have the SNL ack if I am on the paper:

Sandia National Laboratories is a multimission laboratory managed
and operated by National Technology and Engineering Solutions of
Sandia, LLC., a wholly owned subsidiary of Honeywell International,
Inc., for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-NA-0003525.

