\subsection{Descriptive or Prescriptive Semantics}
\label{sub:descriptive}

The high performance community is currently debating the value of 
\emph{descriptive} versus \emph{prescriptive} programming semantics. 
Semantics are descriptive if programming constructs describe the 
computation that should be performed but provide the compiler and 
runtime the flexibility to determine exactly how to perform the 
computation. Programming constructs with prescriptive semantics 
prescribe all details of how to perform the required computation.

Our position is that the debate is misguided since it assumes a binary 
choice between the two types of semantics. However, almost all languages 
have constructs that are descriptive while others are (more) prescriptive. 
Specifically within the HPC community, some claim that OpenACC is descriptive 
while OpenMP is prescriptive~\cite{juckeland2016isc,wolfe16descriptive}. 
While OpenACC provides more descriptive constructs in its most recent version 
than OpenMP does, the \texttt{acc parallel loop} directive is prescriptive 
since sometimes users want to \emph{prescribe} that a loop to be parallelized. 

Alternatively, most OpenMP defaults allow the compiler freedom
to choose details about how the computation is performed. Even
the \texttt{num\_threads} clause of the \texttt{parallel} construct,
which many believe to be among its most prescriptive mechanisms,
allows the compiler and runtime to determine if the number of
threads requested are available. If that many threads are not
available, the compiler and runtime have the flexibility to
determine how many threads to use. So, one may see the issue 
as where to place a language, or even its constructs, on a 
continuum of possible semantics.

More importantly, choosing one place on that continuum is overly
limited and fails to address the overall preference of programmers. 
Specifically, they would prefer that the compiler and runtime would 
always ``do the right thing'' given a description of the computation 
to perform. However, in reality, compilers and runtimes often do not. 
In these instances, programmers prefer to have the ability to override 
their decisions and to prescribe exactly how to perform the computation.

For these reasons, the emerging OpenMP philosophy is to provide
mechanisms that describe the computation to perform and that
prescribe as much or as little as the programmer desires about
how to perform it. As a first step, OpenMP 5.0 will add the 
\texttt{concurrent} construct, which only informs the compiler 
and runtime that a loop nest is easily parallelized. In the 
longer run, we are exploring mechanisms that specify that the 
intent of a clause or a construct is fully descriptive or prescriptive.
