\subsection{Device Extensions}
\label{sub:device_extensions}

While OpenMP introduced support to offload computation regions to target 
devices in version 4.0 and subsequently expanded on the support significantly 
in 4.5, the space is changing quickly. Thus, we have already adopted several
extensions and refinements for OpenMP 5.0. We have adopted changes that
greatly simplify the use of functions in those regions. Further, a new
general mechanism to specify application-specific requirements will enable 
straightforward use of unified memory spaces across devices, Nonetheless, 
we have also adopted a unique deep-copy mechanism that will significantly
improve usability on systems that do not provide unified memory spaces.
Importantly, we expect this deep-copy support will often provide performance 
advantages even on systems that do provide them.

Many offload models, such as CUDA and OpenCL, require function annotations. 
However, OpenMP 5.0 will ease the use of functions on devices by relaxing 
its annotation requirements. OpenMP~5.0 will eliminate the requirement to 
annotate function declarations. Essentially, the compiler must assume that 
a device variant will be available at link time. Also, the compiler must 
automatically generate a device variant for any function with a definition 
in the same translation unit as a call from a \texttt{target} region. 
Essentially, the definition implicitly includes the \texttt{declare target} 
annotation. Because these changes significantly improve usability, many 
compilers have already implemented them and they have allowed entire large 
codebases (particularly in C++ due to the pervasiveness of templates) to 
offload to devices using OpenMP without a single explicit \texttt{declare 
target} directive; other models require hundreds or thousands of annotations 
to compile them at all.

In order to assume coherent memory between the host and a target device, 
the user must assert to the compiler that their code requires that support.
Given this assertion, if the code is run on device without that support, 
it may exhibit unspecified behavior (i.e., the code is broken). Overall,
these assertions represent a contract between the application and the 
compiler, which represents a general mechanism for which unified memory
spaces are just one instance. Thus, OpenMP~5.0 will a new \texttt{requires} 
directive that allows OpenMP to specify a set of rules for a given 
requirement and users to specify that their code accepts those rules for
it to be correct. This directive supports the definition of subsets of 
the OpenMP specification; the first subset will support systems that do 
not require memory to be mapped explicitly into a data environment for 
target devices. Effectively, the user can assume shared memory between
the host and the devices. For example, the code in Figure~\ref{fig:unified} 
is only valid for systems with a unified view of memory. It is non-conforming 
in OpenMP up to 4.5 but will be correct on systems that meet the requirement.
Importantly, the \texttt{requires} directive applies to an entire translation
unit, which offers usability benefits similar to the implicit 
\texttt{declare target} annotations.

\begin{figure}
\begin{minted}{c}
#pragma omp requires \
        unified_shared_memory
struct list {
  void *data;
  struct list *next;
};

void foo() {
  struct list *l = make_linked_list();
#pragma omp target
  {
    struct list *cur;
    while(cur) {
      do_something_with_data(cur->data);
      cur = cur->next;
    }
  }
}
\end{minted}
\caption{How to \texttt{require} a Unified Memory Space}
\label{fig:unified}
\end{figure}

The deep-copy support in OpenMP 5.0 will simplify the use of complex data
structures like the linked list in Figure~\ref{fig:unified} on systems that 
do not provide coherent unified memory. OpenMP 4.5 requires that the user map 
each individual piece of the structure and then assign pointers, to those 
pieces within the data structure on the device either with explicit 
assignments or with further mapping actions. The user often must repeat this 
verbose, complex and error prone code sequence every time an instance of the 
data structure is needed on the device. Instead, \texttt{declare mapper}
directive in OpenMP 5.0 will allow the user to describe how to map a variable 
of that data structure and all of its children or associated data. The user
can then use this definition in a \texttt{map} clause whenever an instance 
of the data structure is needed on the device. Overall, the descriptions in
the \texttt{declare mapper} directive are simpler than the OpenMP 4.5
mechanism and eliminate the repetition. Figure~\ref{fig:mapper} shows the
corresponding example

\begin{figure}
\begin{minted}{cpp}
class Vec {
  size_t len;
  double *data;
public:
  // Normal vector methods
  #pragma omp declare mapper(Vec v) \
    use_by_default \
    map(v, v.data[0:v.len])
};


void foo() {
  Vec v1(100), v2(100);
  fill_vec(v1);
  #pragma omp target teams distribute \
      parallel for \
      map(to:v1) /* explicitly map v1 */
  for(auto i = 0; i < v1.size(); ++i) { 
    v2[i] = v1[i]; // implicitly map v2
  }
  // v2[0-100] == v1[0-100]
}
\end{minted}
\caption{Example usage of \texttt{declare mapper} in a C++ class with private
members}
\label{fig:mapper}
\end{figure}

We plan to refine the deep-copy mechanism further. Specifically, we will 
provide a mechanism that can replace any phase of the mapping process with 
user-defined expressions or functions written in the base language. This 
mechanism, which will provide equivalent functionality to data  serialization 
and deserialization for transmission over a network,  will support mapping 
of arbitrary, complex data structures. Further, it will enable data-dependent 
data transformations that support highly efficient kernel computations. We
are working to include this functionality in OpenMP 5.0 although we may
have to defer its specification for OpenMP 5.1.




